Proportional multi-state multiple-cohort life table model

### 2018-04-09 MSLT/MSLTdoc.Rmd
Pdf figures

---
title: "Proportional multi-state multiple-cohort life table model"
author: "Belen Zapata-Diomedi and Ali Abbas"
date: "26 March 2018"
header-includes:
    - \usepackage{setspace}\onehalfspacing
    - \usepackage{float}
output:
   pdf_document: 
     toc: true
     toc_depth: 3
     number_sections: true
     fig_caption: yes
     fig_width: 7
     fig_height: 6
     df_print: kable
   fontsize: 12 pt
   geometry: margin=1in
bibliography: Bibliography.bib
---

@@

#Global options
knitr::opts_chunk$set(echo = TRUE)

#The work in this document requires the following packages (code, rmarkdown and citing packages):
require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
require(png)
options(knitr.table.format = 'latex')

@@

![Basic ITHIMR infrastructure](structure\Figure1.png)

@@

![Proportional multi-state life-table simplified framework. *The simplied PMST shows the interaction between the life table, disease life table and potential impact fraction (PIF). The PIF calculations by age and sex group are the same as those generated for ITHIM. The PIF (or 1-PIF) modifies incidence of disease, which changes prevalence and mortality (disease specific life table). Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years. Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png)

@@

![Proportional multi-state life-table model. *Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**. The color coding from Data inputs sources to Inputs PMSLT link sources with inputs for the PMSLT. Solid arrows represent final inputs and dashed-arrows represent intermediate inputs that need further processing. Purple coding means a process and green coding represent change in mortality and disability prevalence rates to modify the life table parameters. Black color coding with white color font represent final model outcomes. For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario. **Data inputs** are from four sources: Sythetic population, GBD data (from the Institute of Health Metrics and Evaluations (IHME)), ITHIM calculations and Injuries model. The synthetic population provides population numbers by age-groups and sex which are used in the life table (**PMSLT Mechanisms**) with each age cohort simulated until they die or reach 100 years of age. Population numbers are also used to derive road death rates and road injuries rates from numbers for the baseline and scenario generated by the Injuries model. The difference in road deaths rates modifies all cause-mortatliy in the scenario life table and the change in injuries rates modifies total YLDs rates (hence, healthy life years). Data from GBD IHME includes: total YLD rates per one, disease specific YLDs (numbers) and prevalence to derive disability weights for the included diseases, all cause mortatliy rates per one, population numbers and diseases parameters (incidence, mortality and prevalence). Dismod II is used to derive internally consitent estimates incidence and case fatality rates from GBD IHME data, which are inputs of the disease specific life tables. Case fatality derive from Dismod II is the same for the baseline and scenario (however, could be changed) and incidence is different. For the baseline life tables incidence is as derived from Dismod II and for the scanrio is modified by the PIF (1-PIF). A change in incidence modifies prevalence and mortality in later years. The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png)

>>>>

---
title: "Proportional multi-state multiple-cohort life table model"
author: "Belen Zapata-Diomedi and Ali Abbas"
date: "26 March 2018"
header-includes:
    - \usepackage{setspace}\onehalfspacing
    - \usepackage{float}
    - \usepackage{pdflscape}
    - \newcommand{\blandscape}{\begin{landscape}}
    - \newcommand{\elandscape}{\end{landscape}}
output:
   pdf_document: 
     toc: true
     toc_depth: 3
     number_sections: true

     df_print: kable
   fontsize: 12 pt
   geometry: margin=1in
bibliography: Bibliography.bib
---

@@

#Global options
knitr::opts_chunk$set(echo = TRUE)
options(knitr.table.format = 'latex')
knitr::opts_chunk$set(fig.cap = TRUE, fig.align = "center", fig.pos = "H", fig.scap = TRUE)

#The work in this document requires the following packages (code, rmarkdown and citing packages):
require(dplyr)
require(tidyverse)
require(knitr)
require(kableExtra)
require(citr)
require(png)

@@

![Basic ITHIMR infrastructure](structure\Figure1.png){width=70%, fig.pos="H"}

@@

\newpage
\blandscape

![Proportional multi-state life-table simplified framework. *The simplied PMST shows the interaction between the life table, disease life table and potential impact fraction (PIF). The PIF calculations by age and sex group are the same as those generated for ITHIM. The PIF (or 1-PIF) modifies incidence of disease, which changes prevalence and mortality (disease specific life table). Changes in prevalence and mortality rates from the disease specific life tables feed into the life table by changing all-cause mortality, which in turn changes life years. Change in prevalence of diseases changes total years lived with disability, which in turn modifies health-adjusted life years*](structure\Figure2.png){width=100%, fig.pos="H"}

\elandscape

@@

\newpage
\blandscape

![Proportional multi-state life-table model. *Three sections are presented in Figure 3: **Data input sources**, **Inputs PMSLT** and **PMSLT mechanisms**. The color coding from Data inputs sources to Inputs PMSLT link sources with inputs for the PMSLT. Solid arrows represent final inputs and dashed-arrows represent intermediate inputs that need further processing. Purple coding means a process and green coding represent change in mortality and disability prevalence rates to modify the life table parameters. Black color coding with white color font represent final model outcomes. For both, the life table and disease life table, two sets of each are simulated, one for the baseline and the other for the scenario. **Data inputs** are from four sources: Sythetic population, GBD data (from the Institute of Health Metrics and Evaluations (IHME)), ITHIM calculations and Injuries model. The synthetic population provides population numbers by age-groups and sex which are used in the life table (**PMSLT Mechanisms**) with each age cohort simulated until they die or reach 100 years of age. Population numbers are also used to derive road death rates and road injuries rates from numbers for the baseline and scenario generated by the Injuries model. The difference in road deaths rates modifies all cause-mortatliy in the scenario life table and the change in injuries rates modifies total YLDs rates (hence, healthy life years). Data from GBD IHME includes: total YLD rates per one, disease specific YLDs (numbers) and prevalence to derive disability weights for the included diseases, all cause mortatliy rates per one, population numbers and diseases parameters (incidence, mortality and prevalence). Dismod II is used to derive internally consitent estimates incidence and case fatality rates from GBD IHME data, which are inputs of the disease specific life tables. Case fatality derive from Dismod II is the same for the baseline and scenario (however, could be changed) and incidence is different. For the baseline life tables incidence is as derived from Dismod II and for the scanrio is modified by the PIF (1-PIF). A change in incidence modifies prevalence and mortality in later years. The difference between prevalence and mortality rates between the baseline and scanrio then feed into the scenario life table.* ](structure\Figure3.png){width=70%, hight=50%, fig.pos="H"}

\elandscape
